Displacement effects and implications of Large-Scale Marine Protected Areas

### 2018-09-07 rmd/Manuscript.Rmd
Added new manuscript

---
title: "Displacement effects of Large-Scale MPAs in the Pacific Ocean?"
author: |
  Juan Carlos Villaseñor-Derbez ^1^
  John Lynham ^2^
date: |
  ^1^ Bren School of Environmental Science & Management, University of California, Santa Barbara
  ^2^ Department of Economics, Saunders Hall 532, University of Hawai`i at M~anoa, 2424 Maile Way, Honolulu, HI 96822
abstract: ""
output:
  pdf_document:
    toc: yes
    toc_depth: 4
bibliography: references.bib
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
---

** Last update: `r Sys.Date()`**

```{r, echo = F}
suppressPackageStartupMessages({
  library(tidyverse)
})
```

@@

@@

```{r, echo = F}
pipa_vessels <- read.csv(here::here("data", "vessels_inside_pipa.csv"))

@@

As I mentioned in the call, I have been working with vessels who fished in PIPA as part of my PhD research. In total there are `r length(unique(pipa_vessels$mmsi))`[^3] vessels that have fished within PIPA waters. From these, `r length(unique(n_prev$mmsi))` did so at least once before 2015 and on average exerted `r mean(n_prev$hours)` $\pm$ `r sd(n_prev$hours)` hours each year. A significant number of these vessels fished in PIPA and adjacent waters. We can see where they fished before and track where they went after 2015.

In this section I present some of the preliminary results of regressions similar to what you have conducted for US MNMs using hours as an estimate of fishing effort. First, I explore patterns of fishing effort only inside PIPA (Fig. \ref{fig:blue_para}). Consistent with the blue paradox, we see an increase in the number of vessels, total hours, mean hours, and number of data points between 2014 and 2015. Shortly after that, all measures decrease in 2015[^4].

[^3]: Perhaps some of these vessels only fished there for a couple hours one year. Therefore, I would need to further identify which ones were transient vs local using some cutoff. For now, I use run regressions for all vessels and for vessels that fly the Kiribati flag (n = 12). The true number lies between these two groups, but it provides a starting point.

[^4]: See McDermott et al., (*in review*) for further information

```{r, echo = F, fig.cap = "\\label{fig:blue_para}Monthly hours, number of vessels, and data points observed inside PIPA. These clearly show the blue-paradox effect."}
pipa_vessels %>% 
  group_by(year, month, mmsi) %>% 
  summarize(n_points = sum(n_points),
            hours = sum(total_fishing_hours)) %>% 
  group_by(year, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_points = sum(n_points),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/")),
         post = year >= 2015) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post)) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  geom_point(aes(color = post)) +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

@@

Then, I look at the entire fishing effort of vessels that fished at least once in PIPA before 2015. In other words, I use GFW data to track vessel-level activities inside and outside PIPA before and after 2015. However, all boats included in this case fished at least once in PIPA before 2015 (*i.e.* no control group). First I look at vessels whose `mmsi` code starts with `529`, indicating that they come from Kiribati [^5]. The overall pattern for fishing time (total and mean) show a downward trend that initiated well before 2015 (Fig. \ref{fig:kir_vessels}). The number of vessels as given by unique mmsi numbers shows a maximum near the end of 2014, and then decreases after 2015.

Given the similar trends observed in total fishing hours and mean fishing hours, I calculate total fishing hours at the year-month-vessel level and regress that on a `post` dummy that indicates years before or after 2015. Results of this regression are presented in Table \ref{tab:kir_vessels}. In this table, column (1) presents the regression only against the policy dummy, column (2) includes month fixed effects, column (3) includes month and year fixed effects, and column (4) includes month, year, and vessel fixed effects. All three specifications show negative coefficients. The full model explains 71% of the variance in monthly fishing hours by vessel. This specification has the lowest policy coefficient (labeled `post`), and it is not different from zero at the $\alpha = 0.05$ level.

[^5]: link: [http://www.vtexplorer.com/mmsi-mid-codes-en/](http://www.vtexplorer.com/mmsi-mid-codes-en/)

```{r, echo = F, fig.cap = "\\label{fig:kir_vessels}Fishing hours and number of 12 Kiribati vessels from 2012 - 2016 that at some point fished inside PIPA."}
effort_by_vessel <- read.csv(here::here("data", "pipa_effort_by_vessel.csv")) %>% 
  # filter(between(year, 2014, 2016)) %>%
  filter(mmsi %in% mmsis) %>%
  mutate(mmsi = as.character(mmsi),
         month = as.character(month),
         year = as.character(year))

effort_by_vessel %>% 
  filter(kir) %>% 
  group_by(year, post, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/"))) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post)) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  geom_point(aes(color = post)) +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

```{r, results = "asis", echo = F}
kir_vessel <- effort_by_vessel %>% 
  filter(kir)

m1 <- lm(hours ~ post, data = kir_vessel)
m2 <- lm(hours ~ post + month, data = kir_vessel)
m3 <- lm(hours ~ post + month + year, data = kir_vessel)
m4 <- lm(hours ~ post + month + year + mmsi, data = kir_vessel)

@@

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     type = "latex",
                     omit = c("mmsi", "year", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Year FE", "No", "No", "Yes", "Yes"),
                                      c("Vessel FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:kir_vessels}Fishing hours from GFW for vessels flying Kiribati flag (n = 12). Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.")
```

@@

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Fishing hours and number of vessels by month for all vessels that fished inside PIPA at least once."}
effort_by_vessel %>% 
  group_by(year, post, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/"))) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post)) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  geom_point(aes(color = post)) +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

```{r, results = "asis", echo = F}
m1 <- lm(hours ~ post, data = effort_by_vessel)
m2 <- lm(hours ~ post + month, data = effort_by_vessel)
m3 <- lm(hours ~ post + month + year, data = effort_by_vessel)
m4 <- lm(hours ~ post + month + year + mmsi, data = effort_by_vessel)

@@

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     type = "latex",
                     omit = c("mmsi", "year", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Year FE", "No", "No", "Yes", "Yes"),
                                      c("Vessel FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:all_vessels}Fishing hours from GFW for all vessels (n = 206) that at some point fished in PIPA. Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.")

@@

@@

In order to strengthen this analysis, I would like to identify an appropriate sample of vessels that acccurately represent the PNA. This vessels must meet at least these two points: i) belong to the PNA and ii) Have fished for a significant amount of time inside PIPA, to count as "treated". Ideally, this would get rid of vessels that sporadically fished for insignificant time in PIPA, and are therefore unlikely to be affected. This would include between 12 and `r length(unique(n_prev$mmsi))` vessels (number of Kiribati vessels vs. number of all vessels who fished in PIPA at least once before 2015).

I would also like to explore the possibility of identifying a plausible counterfactual. That way we could do something like a DiD. I think it is possible to identify counterfactual boats. I would start by identifying vessels who meet at least these four criteria: i) fish in near-by regions, ii) have same gear as vessels from PIPA, and iii) having similar length of data, and iv) never fished inside PIPA. Perhaps an additional one would be that they belong to the PNA.

Additionally, run similar regressions on distance traveled, and total vessel hours (not just fishing hours), that could account for the extended "search" time.

```{r, echo = F, fig.height = 3, fig.width = 6, fig.cap = "\\label{fig:eez-hs} Percentage of catches that come from the high-seas vs. EEZs. (data from [Sea Around Us](www.seaaroundus.org))."}
seaaroundus::eezsvshighseas() %>% 
  magrittr::set_colnames(value = c("year", "EEZ", "High-seas")) %>% 
  gather(source, percent, -year) %>% 
  mutate(source = fct_relevel(source, "High-seas")) %>% 
  ggplot2::ggplot(aes(x = year, y = percent, fill = source)) +
  geom_area(alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  geom_hline(yintercept = 94, linetype = "dashed") +
  labs(x = "Year", y = "Percentage", fill = "Source")
```

# Possible contributions

As I understand the efforts made so far, there are two straightforward ways in which I could contribute to this study. For the first one I would help strengthen the analysis of NOAA Observer and GFW data for the US PNMs. The second one would involve conducting the analysis of GFW data for PIPA, and translating the results into implications for Palau. I discuss these two in greater detail in the next paragraphs.

## US Pacific Monuments

In this section I include possible analysis in which I could help. When relevant, I include quotes to provide context. Comments are organized with headings similar to the ones included in the original manuscript. I believe my suggestions would compliment (rather than replace) current analysis, and I acknowledged that you might have already tried some or that due to data constraints not all of them are possible.

### Total revenue

"It is very clear from the figure that total catch of bigeye has not declined as a result of the 2014 or
2016 expansions". Of course, I have not reviewed the data. But is it safe to conclude this? At least through time there isn't an evident decrease, but what would have catches looked like had PMNM not been expanded?

Not sure if the spatial resolution (*i.e.* Catch Blocks) in logbooks allows this comparison, but it would be interesting to see what proportion of catches (or at least sets / hours) originally came from waters inside the PMNM. Perhaps the georeferenced catches from NOAA Observer data can provide insights into this (if not for the entire fleet, at least for the 20% quasi-random subset), or we can use GFW data to see if the proportion of fishing hours inside PMNM roughly corresponds to 10M in revenues (we can test different "catch rates" to account for the fact that catches inside vs outside may not map 1:1). Perhaps this catch efficiency ratio can be extracted from the NOAA Observer data?

On aggregate the fleet brought in more money in 2015 and 1016 than in previous year, but you discuss the possibility that the distributional effects of this might be different. I would assume that vessels who fished the most in the PMNM area were the ones who took the toughest hit. If possible, it might be interesting to birefly look at distributional effects for the discussion.

### Total costs (TC)

So far this section describes the available data, but it it's not clear whether the empirical strategy is fully developed. Just like the datasets, a there isn't a single best analysis that provides the answer, but looking at it from as many was as possible might help paint a full picture of the effects that vLPAs have had.Here I provide some ideas in which we could use the two main datasets (NOAA and GFW data) to indirectly test the claims made by the fleet.

#### TC from NOAA Observer program

You mention that a drawback of this dataset is that you only observe fishing events, and not the entire trip (*i.e.* the search and travel). It might be worth looking into combining this with GFW data. NOAA observer program would provide real fishing events (as opposed to the ones predicted by GFW's neural net), and GFW would provide the entire track. Given that NOAA data cover 20% and GFW only contain 60 boats, we will not get full coverage. If this doesn't yield better data, it would at least help determine if the three-day space between sets is an appropriate way of grouping them into "trips".

Results in your table 3 suggest that there's been a change of ~ 20 miles (positive or negative depending on model specification), and not significant. Perhaps linking this data to GFW would provide full tracks (of a smaller portion of the fleet) to then account for distance between fishing points but also during the search and set activities.

#### TC form GFW Data

We could also test the increased cost claim by looking at the almost 60 vessels and calculate an estimate of fuel consumption (following @sala_2018), which, paired with fuel prices, might provide cost estimates[^1]. Initially this would just be a regression with a policy dummy (and appropriate vessel- and month-level FE's). We could look into GFW data to identify plausible counterfactuals to include  a control group[^2].

[^1]: Of course, there are additional costs (*e.g.* crew, ice, bait, gear), but fuel should represent a large portion of this and should scale with some measures of fishing effort (time and distance). If this is not enough, the previous comparison of catches (sets) that came from PMNM waters before the implementation could provide some insights.

[^2]: GFW increased coverage when they added from more satellites, particularly towards the end of 2016 and throughout 2017, I believe. I think it would be important to use counterfactuals in any GFW-related analyses to account for the expected temporal trend in increased effort.

# Other details

- The introduction states that the main objective is to quantify the economic impacts of vLPAs and compare the cost per hectare to vLPAs on land. I think this would provide an interesting benchmark on cost of conservation per unit area in land vs sea. I would perhaps highlight the part about predicting the possible effects of the Palau MPA.

- I think the novelty of the paper can be highlighted even more. @stevenson_2013 evaluated the consequences of fishing displacement from MPAs in Hawaii (a smaller network and totally different fleet). They show that fishing effort was indeed displaced to places further away from ports representing increased costs, but that CPUE also increased. Overall, they don't find a negative impact on the *perceived* socioeconomic well-being of fishers. However, they do not have plausible counterfacutals and rely on before-after comparisons of perceptions. I think the novelty of this paper is not only the size, resolution, and quality of the data, but the fact that it has a BACI design enabling the identification of the net effect that vLPAs have on fishers. And also the larger contribution to vLPAs (and MPAs in general) literature.
- Recent work by @sala_2018 shows that a large portion of high-seas fishing is only profitable thanks to subsidies. This could compliment @white_2014 and @sumaila_2015 in the introduction.

- "Neither of these datasets is perfect but hopefully combined they will paint a clear picture of what the impact has been". I find this to be a modest (or humble?) and powerful sentence that highlights the efforts made to identify effects from different sources, and that acknowledges possible data limitations. This sentence should be included alongside the "novelty" in the last paragraph of the introduction.

>>>>

---
title: "Displacement effects and implications of Large-Scale Marine Protected Areas"
author: |
  Juan Carlos Villaseñor-Derbez ^1^
  John Lynham ^2^
date: |
  ^1^ Bren School of Environmental Science & Management, University of California, Santa Barbara
  ^2^ Department of Economics, Saunders Hall 532, University of Hawai`i at Manoa
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Literature shows that fishing effort is effectively eliminated in this protected areas upon implementation. The benefits of reducing effort have been largely studied, and include increases in abundance, biomass, and diversity within the bounded regions. In terms of the fishing implication, no-take zones may produce spillover that provide fish for outside areas. The implications of displacement of fishing effort are not yet fully understood. Evidence suggests that fishing effort re-distributes along the borders of the MPA, following an ideal free distribution. Novel data products allow us to track fishing effort at the vessel-level, allowing us to identify changes in behavior at both the fleet and vessel level. This papers evaluates the implications of implementing a LMPA, the Phoenix Islands Protected Area (PIPA). We evaluate changes in fishing effort, and track vessels to see where they go after the implementation of PIPA. Our results show that vessels in the effected region increase fishing effort after the implementation of PIPA. THey are redistributed to X. These findings have major implications for MPA implementation, especially as countries rush to meet the goals of 10% ocean protection as mandated by the SDG."
output:
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 4
    citation_package: natbib
    keep_tex: yes
    number_sections: no
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
bibliography: references.bib
biblio-style: nature
---

*Last update: `r Sys.Date()`*

```{r, echo = F}
suppressPackageStartupMessages({
  library(startR)
  library(tidyverse)
})

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

@@

The rest of the paper is outlined as follows: Section \ref{background} provides further information about PIPA and Palau, section \ref{methods} describes our data and empirical strategy, section \ref{results} presents our results, in section \ref{discussion} we discuss our findings and provide management recommendations.

# Background

@@

```{r, echo = F}
vessel_info_pipa <- read.csv(here::here("data", "vessel_info_mmsi_inside_pipa.csv"), stringsAsFactors = F) %>% 
  mutate(treated = T)

vessel_info_pna <- read.csv(here::here("data", "vessel_info_mmsi_inside_pna.csv"), stringsAsFactors = F) %>% 
  mutate(treated = F)

vessel_info <- rbind(vessel_info_pipa, vessel_info_pna) %>% 
  rename(gear = inferred_label)

pipa_vessels <- read.csv(here::here("data", "vessels_inside_pipa.csv"))

@@

The Global Fishing Watch (GFW) data contain `r length(unique(pipa_vessels$mmsi))` vessels that have fished within PIPA waters [^3]. From these, `r length(mmsis)` did so at least once before 2015.

[^3]: Perhaps some of these vessels only fished there for a couple hours one year. Therefore, I would need to further identify which ones were transient vs constant fishing events using some cutoff. For now, I run regressions for all vessels and for vessels that fly the Kiribati flag (n = 12). The true number of impacted vessels must lie between these two groups, but it provides a starting point.

```{r}
effort_by_vessel <- read.csv(here::here("data", "pipa_effort_by_vessel.csv"), stringsAsFactors = F) %>% 
  rbind(read.csv(here::here("data", "pna_vessels_outside_pipa.csv"), stringsAsFactors = F)) %>% 
  left_join(vessel_info, by = c("mmsi", "year")) %>% 
  filter(gear %in% c("purse_seines", "drifting_longlines")) %>%
  mutate(mmsi = as.character(mmsi),
         month = as.character(month),
         year = as.character(year))

effort_by_vessel %>%  
  group_by(gear, treated, post) %>% 
  summarize(hours = mean(hours, na.rm = T)) %>% 
  spread(post, hours) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treated", "Before", "After", "Change (A / B)"),
               caption = "Mean fishing hours for each gear before and after")
```

```{r}
effort_by_vessel %>% 
  group_by(gear, treated, post, mmsi) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(gear, treated, post) %>% 
  summarize(boats = n()) %>% 
  spread(post, boats) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treated", "Before", "After", "Change (A / B)"),
               caption = "Number of fishing vessels (identified by mmsi) for each gear before and after PIPA implementation.")
```

Then, I look at the entire fishing effort of vessels that fished at least once in PIPA before 2015. In other words, I use GFW data to track vessel-level activities inside and outside PIPA before and after 2015. However, all boats included in this case fished at least once in PIPA before 2015 (*i.e.* no control group). First I look at vessels whose `mmsi` code starts with `529`, indicating that they come from Kiribati [^4]. The overall pattern for fishing time (total and mean) show a downward trend that initiated well before 2015 (Fig. \ref{fig:kir_vessels}). The number of vessels as given by unique mmsi numbers shows a maximum near the end of 2014, and then decreases after 2015.

[^4]: link: [http://www.vtexplorer.com/mmsi-mid-codes-en/](http://www.vtexplorer.com/mmsi-mid-codes-en/)

### To do

- Divide pipa vessels by type of boat (longline vs. purse seine)
- Identify two counterfactuals:
    - A:
        - PNA
        - similar vessel
        - never fished inside PIPA
    -B:
        - Taiwan
        - Never in PNA
        - Similar region?

Has effort increased?

- Change in fishing hours before vs after
- Change in total hours before vs after
- Change in distance before vs after
- Where do PIPA vessels go?
- How many were transient vs local?

Has search time increased?

- Distance between fishing points before vs. after
- Prop hour fishing before vs after
- Time between fishing points before vs. after

### Model specifications

<!-- $$ -->
<!-- H_{ijk} = \alpha + \beta Post_i + \mu_j + \phi_k + \epsilon_{ijk} -->
<!-- $$ -->

<!-- Where $H_{ijk}$ is a measure of effort (*.i.e.* total hours, fishing hours, distance traveled), $\beta$ captures the change for before ($Post_i = 0$) and after ($Post_i = 1$) implementation of PIPA, $mu_j$ captures the month fixed-effects, and $\phi_k$ captures vessel-level fixed-effects. We use this general model and the nested specifications which drop the month and vessel-level fixed effects to test for differences. -->

Our model specification uses a difference-in-differences approach. In this case, we interact the before-after dummy with a treatment dummy. The treatment-control groups follow our definition in the data description. The model takes the form of:

$$
H_{ijk} = \alpha + \beta Post_i \times Treat + \mu_j + \phi_k + \epsilon_{ijk}
$$

In this section I present some of the preliminary results of regressions similar to what you have conducted for US MNMs using hours as an estimate of fishing effort. First, I explore patterns of fishing effort only inside PIPA (Fig. \ref{fig:blue_para}). Consistent with the blue paradox \citep{mcdermott_2018}, we see an increase in the number of vessels, total hours, mean hours, and number of data points between 2014 and 2015. Shortly after that, all measures decrease in 2015.

```{r, echo = F, fig.cap = "\\label{fig:blue_para}Monthly hours, number of vessels, and data points observed inside PIPA. These clearly show the blue-paradox effect."}
# This is just to show the blue paradox, no need to include this figure
pipa_vessels %>% 
  group_by(year, month, mmsi) %>% 
  summarize(n_points = sum(n_points),
            hours = sum(total_fishing_hours)) %>% 
  group_by(year, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_points = sum(n_points),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/")),
         post = year >= 2015) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post)) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  geom_point(aes(color = post)) +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

@@

Given the similar trends observed in total fishing hours and mean fishing hours, I calculate total fishing hours at the year-month-vessel level and regress that on a `post` dummy that indicates years before or after 2015, which is interacted with a `treated` dummy that indicates if vessels fished at least once within PIPA or only in PNA waters. Results of this regression are presented in Table \ref{tab:kir_vessels}. In this table, column (1) presents the regression only against the policy dummy, column (2) includes month fixed effects, column (3) includes month and year fixed effects, and column (4) includes month, year, and vessel fixed effects. All three specifications show negative coefficients. The full model explains 71% of the variance in monthly fishing hours by vessel. This specification has the lowest policy coefficient (labeled `post`), and it is not different from zero at the $\alpha = 0.05$ level.

```{r, echo = F, fig.cap = "\\label{fig:kir_vessels}Fishing hours and number of 12 Kiribati vessels from 2012 - 2016 that at some point fished inside PIPA."}
effort_by_vessel %>% 
  filter(kir, treated) %>% 
  group_by(year, month, mmsi, gear) %>% 
  summarize(hours = sum(hours)) %>% 
  group_by(year, month, gear) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/")),
         post = year >= 2015) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post, gear)) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  geom_point(aes(color = post)) +
  facet_wrap(gear~variable, scales = "free_y", ncol = 3) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

```{r, results = "asis", echo = F}
kir_vessel <- effort_by_vessel %>% 
  filter(kir, gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1)

m1 <- lm(hours ~ post*treated, data = kir_vessel)
m2 <- lm(hours ~ post*treated + month, data = kir_vessel)
m3 <- lm(hours ~ post*treated + month + year, data = kir_vessel)
m4 <- lm(hours ~ post*treated + month + year + mmsi, data = kir_vessel)

@@

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = c("mmsi", "year", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Year FE", "No", "No", "Yes", "Yes"),
                                      c("Vessel FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:kir_vessels}Fishing hours from GFW for vessels flying Kiribati flag (n = 12). Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.")
```

@@

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Fishing hours and number of vessels by month for all vessels that fished inside PIPA at least once."}
effort_by_vessel %>% 
  group_by(year, gear, treated, post, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/"))) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post, gear, treated)) %>% 
  ggplot(aes(x = date, y = value, group = treated)) +
  geom_line(aes(linetype = treated)) +
  geom_point(aes(color = post)) +
  facet_wrap(gear~variable, scales = "free", ncol = 3) +
  theme_bw() +
  scale_color_brewer(palette = "Set1")
```

```{r, results = "asis", echo = F}

model_data <- effort_by_vessel %>% 
  filter(gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1) %>% 
  mutate_at(.vars = vars(month, year, mmsi), .funs = as.character)

m1 <- lm(hours ~ post*treated, data = effort_by_vessel)
m2 <- lm(hours ~ post*treated + month, data = effort_by_vessel)
m3 <- lm(hours ~ post*treated + month + year, data = effort_by_vessel)
m4 <- lm(hours ~ post*treated + month + year + mmsi, data = effort_by_vessel)

@@

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = c("mmsi", "year", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Year FE", "No", "No", "Yes", "Yes"),
                                      c("Vessel FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:all_vessels}Fishing hours from GFW for all vessels (n = 206) that at some point fished in PIPA. Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robuste standard errors.")

@@

```{r, echo = F}
map_df(models, broom::tidy, .id = "model") %>%
  filter(term %in% c("(Intercept)", "postTRUE", "treatedTRUE", "postTRUE:treatedTRUE")) %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error),
                size = 1,
                width = 0,
                color = "red") +
  geom_point(color = "steelblue", size = 3) +
  facet_wrap(~term, scale = "free") +
  theme_bw()
```

# Discusion

@@

<!-- In order to strengthen this analysis, I would like to identify an appropriate sample of vessels that acccurately represent the PNA. This vessels must meet at least these two points: i) belong to the PNA and ii) Have fished for a significant amount of time inside PIPA, to count as "treated". Ideally, this would get rid of vessels that sporadically fished for insignificant time in PIPA, and are therefore unlikely to be affected. This would include between 12 and `r length(unique(n_prev$mmsi))` vessels (number of Kiribati vessels vs. number of all vessels who fished in PIPA at least once before 2015). -->

<!-- I would also like to explore the possibility of identifying a plausible counterfactual. That way we could do something like a DiD. I think it is possible to identify counterfactual boats. I would start by identifying vessels who meet at least these four criteria: i) fish in near-by regions, ii) have same gear as vessels from PIPA, and iii) having similar length of data, and iv) never fished inside PIPA. Perhaps an additional one would be that they belong to the PNA. -->

<!-- Additionally, run similar regressions on distance traveled, and total vessel hours (not just fishing hours), that could account for the extended "search" time. -->

<!-- \clearpage -->

<!-- # Possible contributions -->

<!-- As I understand the efforts made so far, there are two straightforward ways in which I could contribute to this study. For the first one I would help strengthen the analysis of NOAA Observer and GFW data for the US PNMs. The second one would involve conducting the analysis of GFW data for PIPA, and translating the results into implications for Palau. I discuss these two in greater detail in the next paragraphs. -->

<!-- ## US Pacific Monuments -->

<!-- In this section I include possible analysis in which I could help. When relevant, I include quotes to provide context. Comments are organized with headings similar to the ones included in the original manuscript. I believe my suggestions would compliment (rather than replace) current analysis, and I acknowledged that you might have already tried some or that due to data constraints not all of them are possible. -->

<!-- ### Total revenue -->

<!-- "It is very clear from the figure that total catch of bigeye has not declined as a result of the 2014 or -->
<!-- 2016 expansions". Of course, I have not reviewed the data. But is it safe to conclude this? At least through time there isn't an evident decrease, but what would have catches looked like had PMNM not been expanded? -->

<!-- Not sure if the spatial resolution (*i.e.* Catch Blocks) in logbooks allows this comparison, but it would be interesting to see what proportion of catches (or at least sets / hours) originally came from waters inside the PMNM. Perhaps the georeferenced catches from NOAA Observer data can provide insights into this (if not for the entire fleet, at least for the 20% quasi-random subset), or we can use GFW data to see if the proportion of fishing hours inside PMNM roughly corresponds to 10M in revenues (we can test different "catch rates" to account for the fact that catches inside vs outside may not map 1:1). Perhaps this catch efficiency ratio can be extracted from the NOAA Observer data? -->

<!-- On aggregate the fleet brought in more money in 2015 and 1016 than in previous year, but you discuss the possibility that the distributional effects of this might be different. I would assume that vessels who fished the most in the PMNM area were the ones who took the toughest hit. If possible, it might be interesting to birefly look at distributional effects for the discussion. -->

<!-- ### Total costs (TC) -->

<!-- So far this section describes the available data, but it it's not clear whether the empirical strategy is fully developed. Just like the datasets, a there isn't a single best analysis that provides the answer, but looking at it from as many was as possible might help paint a full picture of the effects that vLPAs have had.Here I provide some ideas in which we could use the two main datasets (NOAA and GFW data) to indirectly test the claims made by the fleet. -->

<!-- #### TC from NOAA Observer program -->

<!-- You mention that a drawback of this dataset is that you only observe fishing events, and not the entire trip (*i.e.* the search and travel). It might be worth looking into combining this with GFW data. NOAA observer program would provide real fishing events (as opposed to the ones predicted by GFW's neural net), and GFW would provide the entire track. Given that NOAA data cover 20% and GFW only contain 60 boats, we will not get full coverage. If this doesn't yield better data, it would at least help determine if the three-day space between sets is an appropriate way of grouping them into "trips". -->

<!-- Results in your table 3 suggest that there's been a change of ~ 20 miles (positive or negative depending on model specification), and not significant. Perhaps linking this data to GFW would provide full tracks (of a smaller portion of the fleet) to then account for distance between fishing points but also during the search and set activities. -->

<!-- #### TC form GFW Data -->

<!-- We could also test the increased cost claim by looking at the almost 60 vessels and calculate an estimate of fuel consumption (following @sala_2018), which, paired with fuel prices, might provide cost estimates[^1]. Initially this would just be a regression with a policy dummy (and appropriate vessel- and month-level FE's). We could look into GFW data to identify plausible counterfactuals to include  a control group[^2]. -->

<!-- [^1]: Of course, there are additional costs (*e.g.* crew, ice, bait, gear), but fuel should represent a large portion of this and should scale with some measures of fishing effort (time and distance). If this is not enough, the previous comparison of catches (sets) that came from PMNM waters before the implementation could provide some insights. -->

<!-- [^2]: GFW increased coverage when they added from more satellites, particularly towards the end of 2016 and throughout 2017, I believe. I think it would be important to use counterfactuals in any GFW-related analyses to account for the expected temporal trend in increased effort. -->

<!-- # Other details -->

<!-- - The introduction states that the main objective is to quantify the economic impacts of vLPAs and compare the cost per hectare to vLPAs on land. I think this would provide an interesting benchmark on cost of conservation per unit area in land vs sea. I would perhaps highlight the part about predicting the possible effects of the Palau MPA. -->

<!-- - I think the novelty of the paper can be highlighted even more. @stevenson_2013 evaluated the consequences of fishing displacement from MPAs in Hawaii (a smaller network and totally different fleet). They show that fishing effort was indeed displaced to places further away from ports representing increased costs, but that CPUE also increased. Overall, they don't find a negative impact on the *perceived* socioeconomic well-being of fishers. However, they do not have plausible counterfacutals and rely on before-after comparisons of perceptions. I think the novelty of this paper is not only the size, resolution, and quality of the data, but the fact that it has a BACI design enabling the identification of the net effect that vLPAs have on fishers. And also the larger contribution to vLPAs (and MPAs in general) literature. -->
<!-- - Recent work by @sala_2018 shows that a large portion of high-seas fishing is only profitable thanks to subsidies. This could compliment @white_2014 and @sumaila_2015 in the introduction. -->

<!-- - "Neither of these datasets is perfect but hopefully combined they will paint a clear picture of what the impact has been". I find this to be a modest (or humble?) and powerful sentence that highlights the efforts made to identify effects from different sources, and that acknowledges possible data limitations. This sentence should be included alongside the "novelty" in the last paragraph of the introduction. -->

<!-- - Findings by @smith_2006 show negative effects, but the reserves were only in place for a short period. I believe they talk about this in their discussion. @edgar_2014-UO and @difranco_2016-Xw show how the effects (at least inside vs outside the reserve) can take up to 10 years. -->
